\documentclass[11pt]{article}
\usepackage{geometry}                
\geometry{letterpaper}                 
\usepackage[parfill]{parskip}        
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{epstopdf}
\usepackage{verbatim}
\usepackage{float}
\usepackage{fullpage}
\usepackage{enumerate}
\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\usepackage{gensymb}
\usepackage[T1]{fontenc}
\usepackage[hang,small]{caption}
\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}

\usepackage{listings}
\usepackage{color}
\usepackage{textcomp}
\definecolor{listinggray}{gray}{0.9}
\definecolor{lbcolor}{rgb}{1,1,1}
\lstset{
	backgroundcolor=\color{lbcolor},
	tabsize=4,
	rulecolor=,
	language=matlab,
	basicstyle= \scriptsize,
	upquote=true,
	aboveskip={1.5\baselineskip},
	columns=fixed,
        	showstringspaces=false,
        	extendedchars=true,
        	breaklines=true,
        	prebreak = \raisebox{0ex}[0ex][0ex]{\ensuremath{\hookleftarrow}},
        	frame=single,
        	showtabs=false,
        	showspaces=false,
        	showstringspaces=false,
        	identifierstyle=\ttfamily,
        	keywordstyle=\color[rgb]{0,0,1},
        	commentstyle=\color[rgb]{0.133,0.545,0.133},
        	stringstyle=\color[rgb]{0.627,0.126,0.941},
}


\begin{document}

\section*{Solutions for Discussion 1, 09/11/13}
Compiled by Mai Le
% $x[n]$ can refer to a sequence or a value

\section{Alternate Expressions for Sequences}
Let $x[n] = \begin{cases} \left(\frac{1}{2}\right)^n, & n \text{ nonnegative multiple of 4} \\ -\left(\frac{1}{2}\right)^n, & n \text{ nonnegative multiple of 2, but not a nonnegative multiple of 4} \\ 0, & \text{otherwise}\end{cases}$ \\
Express $x[n]$ mathematically in three different ways.

{\color{blue}
1. $x[n] = \{\underline{1}, 0, \frac{-1}{4}, 0, \frac{1}{16}, 0, \frac{-1}{64},... \}$ \\
\\
2. $x[n] = \delta[n]-\frac{1}{4}\delta[n-2]+\frac{1}{16}\delta[n-4]-\frac{1}{64}\delta[n]+...$ \\ \\
3. $x[n] = \sum\limits_{k=0}^\infty (-1)^k(\frac{1}{4})^k\delta[b-2k]$ \\ \\
4. $x[n] = u[n]cos\left(\frac{\pi}{2}n\right)\left(\frac{1}{2}\right)^n$ }

\section{Nonlinear Systems}
Give an example of a system that is nonlinear but satisfies $\mathcal{T}\{\alpha x[n]\} = \alpha \mathcal{T}\{x[n]\}$ for all sequences $x[n]$ and for all scalars $\alpha \in \mathbb{R}$.

{\color{blue}
$y[n] = \begin{cases} \frac{x^2[n]}{x[n-1]}, & x[n-1] \neq 0 \\
x[n], & x[n-1] = 0 \end{cases} $

Then 
\begin{eqnarray*}\mathcal{T}\{\alpha x[n]\} &=& \begin{cases} \frac{\alpha^2 x^2[n]}{\alpha x[n-1]}, & x[n-1] \neq 0 \\
\alpha x[n], & x[n-1] = 0 \end{cases}  \\
&=& \begin{cases} \frac{\alpha x^2[n]}{x[n-1]}, & x[n-1] \neq 0 \\
\alpha x[n], & x[n-1] = 0 \end{cases} \\
&=& \alpha \begin{cases} \frac{ x^2[n]}{x[n-1]}, & x[n-1] \neq 0 \\
x[n], & x[n-1] = 0 \end{cases} \\
&=& \alpha \mathcal{T}\{x[n]\}
\end{eqnarray*}

but the system fails the addition property:

$\mathcal{T}\{x_1[n]+x_2[n]\} \neq \mathcal{T}\{x_1[n]\} + \mathcal{T}\{x_2[n]\}$

example: Let $x_1[n] = \delta[n]$ and $x_2[n] = \delta[n-1]$. Then $\mathcal{T}\{x_1[n]\} = \delta[n]$ and $\mathcal{T}\{x_2[n]\} = \delta[n-1]$. But $\mathcal{T}\{x_1[n]+x_2[n]\} = \mathcal{T}\{\delta[n]+\delta[n-1]\} = $

%for n=0, 1
%for n=1, 1^2/1 = 1
%for n=2, 0/1 = 0
%
%HMM DIFFERENT EXAMPLE
}

\section{Length of Convolution}
Let $x[n]$ be non-zero only over $N_1 \leq n \leq N_2$ and $h[n]$ be non-zero only over $M_1 \leq n \leq M_2$. Let $y[n]=x[n]*h[n]$. Then $y[n]$ is only non-zero over $L_1 \leq n \leq L_2$. Define $L_1$ and $L_2$ in terms of $N_1$, $N_2$, $M_1$, and $M_2$.

{\color{blue}
For convenience, let $h[n]$ be shorter than $x[n]$. To answer this, we'll envision the "flip and slide", where we flip $h[n]$ and slide it over $x[n]$. Then it goes through three stages: entering, moving across, and exiting.

For $M_1-M_2$ steps, the reversed $h$ is entering into $x[n]$. For $N_1-N_2-(M_1-M_2)-1$ steps, $h[n]$ moves within/across $x[n]$. Finally, $h[n]$ takes $M_1-M_2$ steps to exit $x[n]$. 

The total length, $L_1-L_2 = (M_1-M_2)+(N_1-N_2-(M_1-M_2)-1) + (M_1-M_2) = N_1-N_2+M_1-M_2-1$.

More specifically, $L_1$ is when $h[n]$ is just beginning to enter $x[n]$, so $L_1 = N_1+M_1$, and $L_2$ is when $h[n]$ has finished exiting $x[n]$ so $L_2=N_2+M_2+1$.

}

\section{Distributivity of Convolution}
Prove the distributive property of convolution.

{\color{blue}
\begin{eqnarray*}
x_1[n]*\left(x_2[n]+x_3[n] \right) &=& \sum\limits_{k=-\infty}^\infty x_1[k] \left(x_2[n-k]+x_3[n-k] \right) \text{from definition of convolution} \\
&=& \sum\limits_{k=-\infty}^\infty x_1[k] \left(x_2[n-k]\right)+\sum\limits_{k=-\infty}^\infty x_1[k] \left(x_3[n-k] \right) \\
&=& x_1[n]*x_2[n] + x_1[n]*x_3[n] 
\end{eqnarray*}
}

\section{Computing Discrete Convolution}
Let $y[n]=x[n]*h[n]$. Find an expression for $y[n]$.
\subsection*{(a)}
$x[n] = \begin{cases} 1, & n = -2,0,1 \\ 2, & n = -1\\ 0, & otherwise \end{cases}$
\\
$h[n]=\delta[n]-\delta[n-1]+\delta[n-4]+\delta[n-5]$

{\color{blue}
cry
}

\subsection*{(b)}
$x[n]=u[n+1]-u[n-4]-\delta[n-5]$
\\
$h[n] = \left(u[n+2]-u[n-3]\right)\left(3-|n|\right)$

{\color{blue}
cry
}

\section{Convolution and Signal Sums}
Let $y[n] = x[n]*h[n]$. Prove that $\left( \sum\limits_{n=- \infty }^\infty y[n]\right) = \left(\sum\limits_{n=-\infty}^\infty x[n]\right)\left(\sum\limits_{n=-\infty}^\infty h[n]\right)$.

{\color{blue}
\begin{eqnarray*}
y[n] &=& \sum\limits_{k=-\infty}^\infty h[k]x[n-k] \\
\left(\sum\limits_{n=-\infty}^\infty y[n] \right) &=& \sum\limits_{n=-\infty}^\infty \left( \sum\limits_{k=-\infty}^\infty h[k]x[n-k] \right) \\
 &=& \sum\limits_{k=-\infty}^\infty \left( \sum\limits_{n=-\infty}^\infty h[k]x[n-k] \right) \text{switch order of sums} \\
&=& \sum\limits_{k=-\infty}^\infty h[k] \left(\sum\limits_{n=-\infty}^\infty x[n-k] \right) \text{part in parentheses has same value for all $k$}\\
&=& \left(\sum\limits_{k=-\infty}^\infty h[k]\right) \left(\sum\limits_{m=-\infty}^\infty x[m] \right) \text{change of indexing variable for $x$, $m=n-k$} \\
\end{eqnarray*}
}

\section{Impulse Response of a BIBO Stable System}
Let $h[n]$ be the impulse response of a BIBO stable system. (Remember that impulse responses are only defined for LSI systems, so this system is also LSI.) What must hold true for $h[n]$?

{\color{blue}
Let $x[n]$ be a bounded input with bound $B_x$. Then $|x[n]| \leq B_x$ for all $n$.
\begin{eqnarray*}
|y[n]| &=& \big|\sum\limits_{k=-\infty}^\infty h[k]x[n-k]\big|
\leq  \sum\limits_{k=-\infty}^\infty \big| h[k]x[n-k]\big| \text{ from triangle inequality} \\
\sum\limits_{k=-\infty}^\infty \big| h[k]x[n-k]\big| &\leq & \sum\limits_{k=-\infty}^\infty \big| h[k]\big|\ \big|x[n-k]\big| \leq  \sum\limits_{k=-\infty}^\infty \big| h[k]\big|  B_x = B_x \left(\sum\limits_{k=-\infty}^\infty \big| h[k]\big| \right) = B_y
\end{eqnarray*}

We choose $B_x \left(\sum\limits_{k=-\infty}^\infty \big| h[k]\big| \right)$ as our bound $B_y$. Since $B_x < \infty$, we can ensure $B_y < \infty$ if $\sum\limits_{k=-\infty}^\infty \big| h[k]\big| < \infty$. In other words, if $h[n]$ is absolutely summable (the sum of its absolute values is finite), the output of a bounded input is also bounded. Thus, a system is BIBO stable if its impulse response is absolutely summable. (It is also true that a LSI system with an absolutely summable impulse response is BIBO stable.)
}

\section{Impulse Response of a Causal System}
Let $h[n]$ be the impulse response of a causal system. What must hold true for $h[n]$?

{\color{blue}
\begin{eqnarray*}
y[n] &=& \sum\limits_{k=-\infty}^\infty h[k]x[n-k] \\
&=& \sum\limits_{k=-\infty}^{-1} h[k]x[n-k] \\
&& \quad +\ \ h[0]x[n-0] \\
&& \quad +\ \sum\limits_{k=1}^\infty h[k]x[n-k] \\
&=& \sum\limits_{k=1}^\infty h[-k]x[n+k] \quad \text{future values of $x[n]$} \\
&& \quad +\ \ h[0]x[n-0] \quad \text{present value of $x[n]$}\\
&& \quad +\ \sum\limits_{k=1}^\infty h[k]x[n-k] \quad \text{past values of $x[n]$}\\
\end{eqnarray*}

The first term is a sum of future values of $x[n]$. For the system to be causal, $y[n]$ cannot depend on any of these values- we want each term in that sum to be zero, regardless of the input $x[n]$. We can guarantee that by choosing $h[n] = 0$ for $n < 0$. Thus, a causal system has a causal impulse response. (It is also true that an LSI system with a causal impulse response is a causal system.)
}

\section{Impulse Response of an Invertible System}
A system $\mathcal{T}_1$ is invertible if there exists a system $\mathcal{T}_2$ such that $\mathcal{T}_2\{\mathcal{T}_1\{x[n]\}\} = x[n]$ for all $n \in \mathbb{Z}$. Let $h[n]$ be the impulse response of an invertible system. What must hold true for $h[n]$?

{\color{blue}
For this problem, you can take for granted that the inverse of an LSI system is also LSI.

If $h[n]$ is the impulse response of an invertible system $\mathcal{T}_1$, then there also exists an LSI system $\mathcal{T}_2$ such that $\mathcal{T}_2\{\mathcal{T}_1\{x[n]\}\} = x[n]$. 

\begin{eqnarray*}
\mathcal{T}_2\{\mathcal{T}_1\{x[n]\}\} &=& x[n] \\
\mathcal{T}_2\{h_1[n]*x[n]\} &=& x[n] \\
h_2[n]*(h_1[n]*x[n]) &=& x[n] \text{ where $h_2[n]$ is the impulse response of $\mathcal{T_2}$}\\
(h_2[n]*h_1[n])*x[n]) &=& x[n] \text{ associativity of convolution}\\
(h_2[n]*h_1[n])*x[n]) &=& \delta[n]*x[n] \text{ property of Kronecker delta}\\
\end{eqnarray*}

The Kronecker delta has the property that $\delta[n]*x[n]=x[n]$ for all $n$. It is also the only discrete signal with that property. Therefore, $(h_2[n]*h_1[n])=\delta[n]$. In conclusion, if a system is invertible with impulse response $h_1[n]$, then there exists some other sequence $h_2[n]$ such that $(h_2[n]*h_1[n])=\delta[n]$, and $h_2[n]$ is the impulse response of the inverse system.
}

\section{Eigensequences}
A sequence $x[n]$ is an eigensequence of a system $\mathcal{T}$ if $\mathcal{T}\{x[n]\}=\lambda x[n]$ for some non-zero scalar $\lambda \in \mathbb{C}$. What are the eigensequences for the following systems?
\subsection{$\mathcal{T}\{x[n]\}=3x[n]$}

{\color{blue}
Any sequence is an eigensequence of this system!
}

\subsection{$\mathcal{T}\{x[n]\}=x[n]u[n]$}
{\color{blue}
Any causal sequence is an eigensequence of this system.
}

\subsection{causal moving average: $\mathcal{T}\{x[n]\}=\frac{1}{M}\sum\limits_{k=0}^{M-1} x[n-k]$}

{\color{blue}
Any constant sequence is an eigensequence of this system. Complex exponentials of the form $x[n] = Ae^{j\omega n}$ are also eigenfunctions (because this is an LSI system).
}

\subsection{general LSI system: $\mathcal{T}\{x[n]\}=\sum\limits_{k=-\infty}^\infty h[k]x[n-k]$}

{\color{blue}
We showed this in lecture 3.

Any sequence of the form $x[n] = Ae^{j\omega n}$ is an eigensequence:
\begin{eqnarray*}
\mathcal{T}\{Ae^{j\omega n}\}&=&\sum\limits_{k=-\infty}^\infty h[k]Ae^{j\omega (n-k)} \\
&=& Ae^{j\omega n} \left(\sum\limits_{k=-\infty}^\infty h[k]e^{-j\omega k} \right) \\
&=& Ae^{j\omega n} \lambda 
\end{eqnarray*}

Since $\sum\limits_{k=-\infty}^\infty h[k]e^{-j\omega k}$ is not a function of $n$, we can rewrite is as a constant $\lambda$ and see that the output is a scaled version of the input.
}

\section{Geometric Basis for Sequences}
Consider the signal $\gamma[n]=a^nu[n]$ for $0<a<1$. 

\subsection*{(a)} 
Show that any sequence $x[n]$ can be decomposed as $x[n]=\sum\limits_{n=-\infty}^\infty c_k \gamma[n-k]$ and express $c_k$ in terms of $x[n]$. 

\subsection*{(b)} 
Use the properties of linearity and time invariance to express the output $y[n] = \mathcal{T}\{x[n]\}$ in terms of the input $x[n]$ and the signal $g[n]=\mathcal{T}\{y[n]\}$, where $\mathcal{T}$ is an LTI system.

\subsection*{(c)} 
Express the impulse response $h[n]=\mathcal{T}\{\delta[n]\}$ in terms of $g[n]$.

\section{Steady State of Stable Systems}
Let $\mathcal{T}$ be an LTI, BIBO stable system. Show if x[n] is bounded and tends to a constant, the corresponding output, $y[n]$, will also tend to a constant.

{\color{blue}
If $x[n]$ tends towards a constant, then $\lim\limits_{n\rightarrow \infty} x[n] = c_x < \infty$. Since the system is LSI, we also know it has an impulse response $h[n]$. Since it is BIBO stable, we also know from Problem 7 that $\sum\limits_{n=-\infty}^\infty \big| h[n]\big| < \infty$.

\[
\lim\limits_{n\rightarrow \infty} y[n] = \lim\limits_{n\rightarrow \infty} \sum\limits_{k=-\infty}^\infty h[k]x[n-k] = \sum\limits_{k=-\infty}^\infty  \lim\limits_{n\rightarrow \infty} h[k]x[n-k] = \sum\limits_{k=-\infty}^\infty h[k] \left( \lim\limits_{n\rightarrow \infty} x[n-k] \right) =c_y
\]

Since $\lim\limits_{n\rightarrow \infty} x[n-k] = \lim\limits_{n\rightarrow \infty} x[n] = c_x$ and $\sum\limits_{k=-\infty}^\infty h[k] \leq \sum\limits_{k=-\infty}^\infty |h[k]| < \infty$, their product, called $c_y$ is also finite. Thus, the corresponding output $y[n]$ also tends toward a constant.
}

\end{document}