\documentclass[11pt]{article}
\usepackage{geometry}                
\geometry{letterpaper}                 
\usepackage[parfill]{parskip}        
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{epstopdf}
\usepackage{verbatim}
\usepackage{float}
\usepackage{fullpage}
\usepackage{enumerate}
\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\usepackage{gensymb}
\usepackage[T1]{fontenc}
\usepackage[hang,small]{caption}
\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}

\usepackage{listings}
\usepackage{color}
\usepackage{textcomp}
\definecolor{listinggray}{gray}{0.9}
\definecolor{lbcolor}{rgb}{1,1,1}
\lstset{
	backgroundcolor=\color{lbcolor},
	tabsize=4,
	rulecolor=,
	language=matlab,
	basicstyle= \scriptsize,
	upquote=true,
	aboveskip={1.5\baselineskip},
	columns=fixed,
        	showstringspaces=false,
        	extendedchars=true,
        	breaklines=true,
        	prebreak = \raisebox{0ex}[0ex][0ex]{\ensuremath{\hookleftarrow}},
        	frame=single,
        	showtabs=false,
        	showspaces=false,
        	showstringspaces=false,
        	identifierstyle=\ttfamily,
        	keywordstyle=\color[rgb]{0,0,1},
        	commentstyle=\color[rgb]{0.133,0.545,0.133},
        	stringstyle=\color[rgb]{0.627,0.126,0.941},
}


\begin{document}

\section*{Problems for Discussion 1, 09/11/13}
Compiled by Mai Le
% $x[n]$ can refer to a sequence or a value

\section{Alternate Expressions for Sequences}
Let $x[n] = \begin{cases} \left(\frac{1}{2}\right)^n, & n \text{ nonnegative multiple of 4} \\ -\left(\frac{1}{2}\right)^n, & n \text{ nonnegative multiple of 2, but not a nonnegative multiple of 4} \\ 0, & \text{otherwise}\end{cases}$ \\
Express $x[n]$ mathematically in three different ways

\section{Nonlinear Systems}
Give an example of a system that is nonlinear but satisfies $\mathcal{T}\{\alpha x[n]\} = \alpha \mathcal{T}\{x[n]\}$ for all sequences $x[n]$ and for all scalars $\alpha \in \mathbb{R}$.

\section{Length of Convolution}
Let $x[n]$ be non-zero only over $N_1 \leq n \leq N_2$ and $h[n]$ be non-zero only over $M_1 \leq n \leq M_2$. Let $y[n]=x[n]*h[n]$. Then $y[n]$ is only non-zero over $L_1 \leq n \leq L_2$. Define $L_1$ and $L_2$ in terms of $N_1$, $N_2$, $M_1$, and $M_2$.

\section{Distributivity of Convolution}
Prove the distributive property of convolution.

\section{Computing Discrete Convolution}
Let $y[n]=x[n]*h[n]$. Find an expression for $y[n]$.
\subsection*{(a)}
$x[n] = \begin{cases} 1, & n = -2,0,1 \\ 2, & n = -1\\ 0, & otherwise \end{cases}$
\\
$h[n]=\delta[n]-\delta[n-1]+\delta[n-4]+\delta[n-5]$
\subsection*{(b)}
$x[n]=u[n+1]-u[n-4]-\delta[n-5]$
\\
$h[n] = \left(u[n+2]-u[n-3]\right)\left(3-|n|\right)$

\section{Convolution and Signal Energy}
Let $y[n] = x[n]*h[n]$. Prove that $\left( \sum\limits_{n=- \infty }^\infty y[n]\right) = \left(\sum\limits_{n=-\infty}^\infty x[n]\right)\left(\sum\limits_{n=-\infty}^\infty h[n]\right)$.


\section{Impulse Response of a BIBO Stable System}
Let $h[n]$ be the impulse response of a BIBO stable system. (Remember that impulse responses are only defined for LSI systems, so this system is also LSI.) What must hold true for $h[n]$?

\section{Impulse Response of a Causal System}
Let $h[n]$ be the impulse response of a causal system. What must hold true for $h[n]$?

\section{Impulse Response of an Invertible System}
A system $\mathcal{T}_1$ is invertible if there exists a system $\mathcal{T}_2$ such that $\mathcal{T}_2\{\mathcal{T}_1\{x[n]\}\} = x[n]$ for all $n \in \mathbb{Z}$. Let $h[n]$ be the impulse response of an invertible system. What must hold true for $h[n]$?

\section{Eigensequences}
A sequence $x[n]$ is an eigensequence of a system $\mathcal{T}$ if $\mathcal{T}\{x[n]\}=\lambda x[n]$ for some scalar $\lambda \in \mathbb{C}$. What are the eigensequences for the following systems?
\subsection{$\mathcal{T}\{x[n]\}=3x[n]$}

\subsection{$\mathcal{T}\{x[n]\}=x[n]u[n]$}

\subsection{causal moving average: $\mathcal{T}\{x[n]\}=\frac{1}{M}\sum\limits_{k=0}^{M-1} x[n-k]$}

\subsection{general LSI system: $\mathcal{T}\{x[n]\}=\sum\limits_{k=-\infty}^\infty h[k]x[n-k]$}

\section{Geometric Basis for Sequences}
Consider the signal $\gamma[n]=a^nu[n]$ for $0<a<1$. 

\subsection*{(a)} 
Show that any sequence $x[n]$ can be decomposed as $x[n]=\sum\limits_{n=-\infty}^\infty c_k \gamma[n-k]$ and express $c_k$ in terms of $x[n]$. 

\subsection*{(b)} 
Use the properties of linearity and time invariance to express the output $y[n] = \mathcal{T}\{x[n]\}$ in terms of the input $x[n]$ and the signal $g[n]=\mathcal{T}\{y[n]\}$, where $\mathcal{T}$ is an LTI system.

\subsection*{(c)} 
Express the impulse response $h[n]=\mathcal{T}\{\delta[n]\}$ in terms of $g[n]$.

\section{Steady State of Stable Systems}
Let $\mathcal{T}$ be an LTI, BIBO stable system. Show if x[n] is bounded and tends to a constant, the corresponding output, $y[n]$, will also tend to a constant.

\end{document}